---
title: "DE Table and UpSets"
output: html_notebook
---

```{r,message = FALSE}
library(broom)
library(UpSetR)
library(tidyverse)
library(limma)
library(DT)
library(ReporteRs)
load('~/git/unified_gene_expression/data/big_six_DE_correct.Rdata')
load('~/git/unified_gene_expression/data/limma_voom_DE_all_by_all.Rdata')
source('~/git/unified_gene_expression/scripts/GO_enrichment.R')
source('~/git/unified_gene_expression/scripts/GO_term_finder.R')
# for GO enrichement
background_genes <- topTable(big_six[[1]], number=30000) %>% rownames_to_column('Gene') %>% dplyr::select(Gene)
options("ReporteRs-fontsize"=8, "ReporteRs-default-font"="Monaco")
```

Overall counts up and down for the six tests
```{r, message = FALSE}
test1_summary <- summary(decideTests(big_six[[1]], lfc = 1, p.value = 0.01, adjust.method = 'fdr')) %>% tidy() %>% mutate(Comparison = '1')
test2_summary <- summary(decideTests(big_six[[2]], lfc = 1, p.value = 0.01, adjust.method = 'fdr')) %>% tidy() %>% mutate(Comparison = '2')
test3_summary <- summary(decideTests(big_six[[3]], lfc = 1, p.value = 0.01, adjust.method = 'fdr')) %>% tidy() %>% mutate(Comparison = '3')
test4_summary <- summary(decideTests(big_six[[4]], lfc = 1, p.value = 0.01, adjust.method = 'fdr')) %>% tidy() %>% mutate(Comparison = '4')
test5_summary <- summary(decideTests(big_six[[5]], lfc = 1, p.value = 0.01, adjust.method = 'fdr')) %>% tidy() %>% mutate(Comparison = '5')
test6_summary <- summary(decideTests(big_six[[6]], lfc = 1, p.value = 0.01, adjust.method = 'fdr')) %>% tidy() %>% mutate(Comparison = '6')

test_summaries <- rbind(test1_summary,test2_summary,test3_summary,test4_summary,test5_summary,test6_summary)
colnames(test_summaries) <- c('LogFC','Comparison','Count','Comparison Set')

test_summaries <- test_summaries %>% 
  filter(LogFC!=0) %>% 
  dplyr::select(-`Comparison Set`) %>% 
  spread(Comparison, Count) %>% t() %>% 
  data.frame() %>% 
  rownames_to_column('Comparisons') %>% 
  dplyr::slice(-1) %>% 
  separate(Comparisons, c('Comparison','Base'), sep='_vs_') %>% 
  mutate(Comparison=gsub('\\.',' (',Comparison), Comparison=gsub('$',')', Comparison), Comparison=gsub('cell',' cell', Comparison)) %>% 
  mutate(Base=gsub('\\.',' (',Base), Base=gsub('$',')', Base), Base=gsub('cell',' cell', Base))

colnames(test_summaries)<- c('Comparison','Reference', 'LogFC < -1', 'LogFC > 1')
```

Format table 
http://davidgohel.github.io/ReporteRs/articles/flextable_examples.html
```{r}
no_border = borderProperties( width = 0 )
big_border = borderProperties( width = 2 )
std_border = borderProperties( width = 1 )
    
the_table <- test_summaries %>% 
  mutate(Test=c('A','A','A','B','B','B','C','C','C','D','E','F','F','F')) %>% 
  mutate(Comparison = paste(Comparison, Reference, sep = ' vs ')) %>% 
  dplyr::select(Test, Comparison, `LogFC < -1`, `LogFC > 1`)

flex_table <- the_table %>% FlexTable(body.cell.props = cellProperties( padding = 2 ), 
                        header.par.props = parProperties(text.align = "center" ), 
                        body.par.props = parProperties(text.align = 'left'),
                        header.columns = FALSE) %>% 
  spanFlexTableRows(.,j="Test",runs=as.character(the_table$Test)) %>%
  addHeaderRow(value = c("", "Log Fold Change"), colspan = c( 2, 2 ) ) %>% 
  addHeaderRow(value = c('Test', 'Comparison', '< -1', '> 1')) %>% 
  setFlexTableWidths(widths = c(1,5, 1, 1) )  %>% 
  setFlexTableBorders(footer = TRUE,
                      inner.vertical = no_border, inner.horizontal = std_border,
                      outer.vertical = no_border, outer.horizontal = big_border )

flex_table[2,1:2, to="header", side = "top"] = no_border
flex_table
#doc = docx()
#doc = addFlexTable(doc, flex_table)
#writeDoc(doc,file='~/Desktop/table class.docx')
```

```{r, message = FALSE}
set_maker <- function(a,b,c){
  a_set <- setdiff(a,c(b,c))
  b_set <- setdiff(b,c(a,c))
  c_set <- setdiff(c,c(a,b))
  abc <- Reduce(intersect,list(a,b,c))
  ab <- setdiff(intersect(a,b),c)
  ac <- setdiff(intersect(a,c),b)
  bc <- setdiff(intersect(b,c),a)
  list(a=a_set,b=b_set,c=c_set,ab=ab,ac=ac,bc=bc,abc=abc)
}
```

Function to return gene list that are abs(logFC) > 1 and adj.P.Val < 0.05 for the different contrasts
```{r, messag= FALSE}
# all sig
contrast_gene_lists <- function(limma_data){
  vector_of_comparisons <- colnames(limma_data)
  print(vector_of_comparisons)
  out_list <- list()
  for (i in 1:length(vector_of_comparisons)){
    stats <-  topTable(limma_data, coef=i, number = 30000, adjust.method = 'fdr', p.value = 0.05)
    if(nrow(stats)==0){
      out_list[vector_of_comparisons[i]] <- list('')
      next}
    stats_cut_down <- stats[abs(stats[,'logFC']) > 1,]
    out_list[vector_of_comparisons[i]] <- list(row.names(stats_cut_down))
    }
  out_list
}
# just up vs 'control' (right side of contrast)
contrast_UPgene_lists <- function(limma_data){
  vector_of_comparisons <- colnames(limma_data)
  out_list <- list()
  for (i in 1:length(vector_of_comparisons)){
    stats <-  topTable(limma_data, coef=i, number = 30000, adjust.method = 'fdr', p.value = 0.05)
    if(nrow(stats)==0){
      out_list[vector_of_comparisons[i]] <- list('')
      next}
    stats_cut_down <- stats[stats[,'logFC'] > 1,]
    out_list[vector_of_comparisons[i]] <- list(row.names(stats_cut_down))
    }
  out_list
}
# just down vs 'control' 
contrast_DOWNgene_lists <- function(limma_data){
  vector_of_comparisons <- colnames(limma_data)
  out_list <- list()
  for (i in 1:length(vector_of_comparisons)){
    stats <-  topTable(limma_data, coef=i, number = 30000, adjust.method = 'fdr', p.value = 0.05)
    if(nrow(stats)==0){
      out_list[vector_of_comparisons[i]] <- list('')
      next}
    stats_cut_down <- stats[stats[,'logFC'] < -1,]
    out_list[vector_of_comparisons[i]] <- list(row.names(stats_cut_down))
    }
  out_list
}
```

# 1. Cornea vs Retina vs RPE (all adult)
```{r, message = FALSE, fig.width=3, fig.height=2.5}

colnames(big_six[[1]])

test1_contrast_genes <- contrast_gene_lists(big_six[[1]])
# test1_comps <- list('Cornea vs Retina' = test1_contrast_genes$Cornea.adult_vs_Retina.adult, 
#                    'Cornea vs RPE' = test1_contrast_genes$Cornea.adult_vs_RPE.adult, 
#                    'Retina vs RPE' = test1_contrast_genes$Retina.adult_vs_RPE.adult)
# upset(fromList(test1_comps),mainbar.y.max = 3250, keep.order = T)


test1_contrast_genes <- contrast_UPgene_lists(big_six[[1]])
test1_comps <- list('Cornea vs Retina' = test1_contrast_genes$Cornea.adult_vs_Retina.adult, 
                   'Cornea vs RPE' = test1_contrast_genes$Cornea.adult_vs_RPE.adult, 
                   'Retina vs RPE' = test1_contrast_genes$Retina.adult_vs_RPE.adult)
upset(fromList(test1_comps),sets = rev(names(test1_comps)), mainbar.y.max = 3000, keep.order = T, sets.bar.color = '#3eb9c4', matrix.color = '#3eb9c4', main.bar.color = '#3eb9c4', set_size.angles = 45)


test1_contrast_genes <- contrast_DOWNgene_lists(big_six[[1]])
test1_comps <- list('Cornea vs Retina' = test1_contrast_genes$Cornea.adult_vs_Retina.adult, 
                   'Cornea vs RPE' = test1_contrast_genes$Cornea.adult_vs_RPE.adult, 
                   'Retina vs RPE' = test1_contrast_genes$Retina.adult_vs_RPE.adult)
upset(fromList(test1_comps),sets = rev(names(test1_comps)),mainbar.y.max = 3000, keep.order = T, sets.bar.color = '#ef6d62', matrix.color = '#ef6d62', main.bar.color = '#ef6d62', set_size.angles = 45)
```

# 2. ESC vs RPE vs Retina (all Stem cell lines, non immortalized)
```{r, message = FALSE, fig.width=3, fig.height=2.5}
colnames(big_six[[2]])
# test2_contrast_genes <- contrast_gene_lists(big_six[[2]])
# test2_comps <- list('ESC vs Retina (stem cell)' = test2_contrast_genes$ESC.stemcell_vs_Retina.stemcell, 
#                    'ESC vs RPE (stem cell)' = test2_contrast_genes$ESC.stemcell_vs_RPE.stemcell, 
#                    'Retina (stem cell) vs RPE (stem cell)' = test2_contrast_genes$Retina.stemcell_vs_RPE.stemcell)
# upset(fromList(test2_comps), keep.order = T, mainbar.y.max = 3000)

test2_contrast_genes <- contrast_UPgene_lists(big_six[[2]])
test2_comps <- list('ESC vs Retina (stem cell)' = test2_contrast_genes$ESC.stemcell_vs_Retina.stemcell, 
                   'ESC vs RPE (stem cell)' = test2_contrast_genes$ESC.stemcell_vs_RPE.stemcell, 
                   'Retina (stem cell) vs RPE (stem cell)' = test2_contrast_genes$Retina.stemcell_vs_RPE.stemcell)
upset(fromList(test2_comps),sets = rev(names(test2_comps)), keep.order = T, mainbar.y.max = 2500, sets.bar.color = '#3eb9c4', matrix.color = '#3eb9c4', main.bar.color = '#3eb9c4', set_size.angles = 45)

test2_contrast_genes <- contrast_DOWNgene_lists(big_six[[2]])
test2_comps <- list('ESC vs Retina (stem cell)' = test2_contrast_genes$ESC.stemcell_vs_Retina.stemcell, 
                   'ESC vs RPE (stem cell)' = test2_contrast_genes$ESC.stemcell_vs_RPE.stemcell, 
                   'Retina (stem cell) vs RPE (stem cell)' = test2_contrast_genes$Retina.stemcell_vs_RPE.stemcell)
upset(fromList(test2_comps), sets = rev(names(test2_comps)),keep.order = T, mainbar.y.max = 2500, sets.bar.color = '#ef6d62', matrix.color = '#ef6d62', main.bar.color = '#ef6d62', set_size.angles = 45)
```

# 3. Adult RPE vs Fetal RPE vs Cell Line RPE
 RPE (stem cell) vs RPE (fetal) - coef 1
	RPE (stem cell) vs RPE (adult) - coef 2
	RPE (fetal) vs RPE (adult) - (coef 3)
```{r, message = FALSE}

test3_contrast_genes <- contrast_UPgene_lists(big_six[[3]])
test3_comps <- list('RPE (stem cell) vs RPE (fetal)' = test3_contrast_genes$RPE.stemcell_vs_RPE.fetal, 
                   'RPE (stem cell) vs RPE (adult)' = test3_contrast_genes$RPE.stemcell_vs_RPE.adult, 
                   'RPE (fetal) vs RPE (adult)' = test3_contrast_genes$RPE.fetal_vs_RPE.adult)
upset(fromList(test3_comps),sets = rev(names(test3_comps)), keep.order = T, mainbar.y.max = 3000, sets.bar.color = '#3eb9c4', matrix.color = '#3eb9c4', main.bar.color = '#3eb9c4', set_size.angles = 45)

test3_contrast_genes <- contrast_DOWNgene_lists(big_six[[3]])
test3_comps <- list('RPE (stem cell) vs RPE (fetal)' = test3_contrast_genes$RPE.stemcell_vs_RPE.fetal, 
                   'RPE (stem cell) vs RPE (adult)' = test3_contrast_genes$RPE.stemcell_vs_RPE.adult, 
                   'RPE (fetal) vs RPE (adult)' = test3_contrast_genes$RPE.fetal_vs_RPE.adult)
upset(fromList(test3_comps), sets = rev(names(test3_comps)),keep.order = T, mainbar.y.max = 3000, sets.bar.color = '#ef6d62', matrix.color = '#ef6d62', main.bar.color = '#ef6d62', set_size.angles = 45)
```

# 6. Adult Cornea vs Fetal Cornea vs Cell Line Cornea
```{r, message= FALSE}
colnames(big_six[[6]])

test6_contrast_genes <- contrast_UPgene_lists(big_six[[6]])
test6_comps <- list('Cornea vs Cornea (fetal)' = test6_contrast_genes$Cornea.adult_vs_Cornea.fetal,
                   'Cornea vs Cornea (cell line)' = test6_contrast_genes$Cornea.adult_vs_Cornea.immortalizedcell, 
                   'Cornea (fetal) vs Cornea (cell line)' = test6_contrast_genes$Cornea.fetal_vs_Cornea.immortalizedcell)
upset(fromList(test6_comps),sets = rev(names(test6_comps)), keep.order = T, mainbar.y.max = 1750, sets.bar.color = '#3eb9c4', matrix.color = '#3eb9c4', main.bar.color = '#3eb9c4', set_size.angles = 45)

test6_contrast_genes <- contrast_DOWNgene_lists(big_six[[6]])
test6_comps <- list('Cornea vs Cornea (fetal)' = test6_contrast_genes$Cornea.adult_vs_Cornea.fetal,
                   'Cornea vs Cornea (cell line)' = test6_contrast_genes$Cornea.adult_vs_Cornea.immortalizedcell, 
                   'Cornea (fetal) vs Cornea (cell line)' = test6_contrast_genes$Cornea.fetal_vs_Cornea.immortalizedcell)
upset(fromList(test6_comps), sets = rev(names(test6_comps)),keep.order = T, mainbar.y.max = 1750, sets.bar.color = '#ef6d62', matrix.color = '#ef6d62', main.bar.color = '#ef6d62', set_size.angles = 45)
```
