---
title: "Figure 3"
output: html_notebook
---
```{r,message = FALSE}
library(broom)
library(UpSetR)
library(tidyverse)
```

Overall counts up and down for the six tests
```{r, message = FALSE}
test1_summary <- summary(decideTests(big_six[[1]], lfc = 1, p.value = 0.05, adjust.method = 'fdr')) %>% tidy() %>% mutate(Comparison = '1')
test2_summary <- summary(decideTests(big_six[[2]], lfc = 1, p.value = 0.05, adjust.method = 'fdr')) %>% tidy() %>% mutate(Comparison = '2')
test3_summary <- summary(decideTests(big_six[[3]], lfc = 1, p.value = 0.05, adjust.method = 'fdr')) %>% tidy() %>% mutate(Comparison = '3')
test4_summary <- summary(decideTests(big_six[[4]], lfc = 1, p.value = 0.05, adjust.method = 'fdr')) %>% tidy() %>% mutate(Comparison = '4')
test5_summary <- summary(decideTests(big_six[[5]], lfc = 1, p.value = 0.05, adjust.method = 'fdr')) %>% tidy() %>% mutate(Comparison = '5')
test6_summary <- summary(decideTests(big_six[[6]], lfc = 1, p.value = 0.05, adjust.method = 'fdr')) %>% tidy() %>% mutate(Comparison = '6')

test_summaries <- rbind(test1_summary,test2_summary,test3_summary,test4_summary,test5_summary,test6_summary)
colnames(test_summaries) <- c('LogFC','Comparison','Count','Comparison Set')

test_summaries <- test_summaries %>% 
  filter(LogFC!=0) %>% 
  select(-`Comparison Set`) %>% 
  spread(Comparison, Count) %>% t() %>% 
  data.frame() %>% 
  rownames_to_column('Comparisons') %>% 
  slice(-1) %>% 
  separate(Comparisons, c('Comparison','Base'), sep='_vs_') %>% 
  mutate(Comparison=gsub('\\.',' (',Comparison), Comparison=gsub('$',')', Comparison), Comparison=gsub('cell',' cell', Comparison)) %>% 
  mutate(Base=gsub('\\.',' (',Base), Base=gsub('$',')', Base), Base=gsub('cell',' cell', Base))

colnames(test_summaries)<- c('Comparison','Reference', 'LogFC < -1', 'LogFC > 1')

test_summaries %>% mutate(`Test #`=c(1,1,1,2,2,2,3,3,3,4,5,6,6,6))

set_maker <- function(a,b,c){
  a_set <- setdiff(a,c(b,c))
  b_set <- setdiff(b,c(a,c))
  c_set <- setdiff(c,c(a,b))
  abc <- Reduce(intersect,list(a,b,c))
  ab <- setdiff(intersect(a,b),c)
  ac <- setdiff(intersect(a,c),b)
  bc <- setdiff(intersect(b,c),a)
  list(a_set,b_set,c_set,ab,ac,bc,abc)
}

for (i in 1:6){print(colnames(big_six[[i]]))}

cornea_vs_retina <- topTableF(big_six[[1]], number = 30000, adjust.method = 'fdr') %>% 
  rownames_to_column('Gene') %>% 
  filter(abs(Cornea.adult_vs_Retina.adult) > 1, adj.P.Val < 0.05) %>% 
  .[['Gene']]
cornea_vs_rpe <- topTableF(big_six[[1]], number = 30000, adjust.method = 'fdr') %>% 
  rownames_to_column('Gene') %>% 
  filter(abs(Cornea.adult_vs_RPE.adult) > 1, adj.P.Val < 0.05) %>% 
  .[['Gene']]
retina_vs_rpe <- topTableF(big_six[[1]], number = 30000, adjust.method = 'fdr') %>% 
  rownames_to_column('Gene') %>% 
  filter(abs(Retina.adult_vs_RPE.adult) > 1, adj.P.Val < 0.05) %>% 
  .[['Gene']]

all_sets <- set_maker(cornea_vs_retina, cornea_vs_rpe,retina_vs_rpe)
all_sets[[1]]

test_comps <- list('Cornea vs Retina' = cornea_vs_retina, 'Cornea vs RPE' = cornea_vs_rpe, 'Retina vs RPE' = retina_vs_rpe)
upset(fromList(test_comps))

```


Tabulate numbers for boxplot (table?)

```{r, message = FALSE}
comparisons <- colnames(efit$contrasts)
#pretty up the comparison names
comparisons_p <- gsub('\\.',' (', comparisons) %>% gsub('_v',') v',.) %>% gsub('s_','s ',.) %>% gsub('$',')',.)

all_counts <- c()
for (i in seq(1,length(comparisons_p))){
  up_count<- topTable(efit, coef=i, adjust='fdr', number=30000) %>% filter(logFC > 1, adj.P.Val<0.05) %>% nrow()
  up_count <- c(comparisons_p[i], '> 1', up_count)
  down_count <- topTable(efit, coef=i, adjust='fdr', number=30000) %>% filter(logFC < -1, adj.P.Val<0.05) %>% nrow()
  down_count <- c(comparisons_p[i], '< -1', down_count)
  
  all_counts <- rbind(all_counts, up_count, down_count)
}
row.names(all_counts) <- seq(1,nrow(all_counts))
colnames(all_counts) <- c('VS','logFC','Count')
all_counts <- all_counts %>% 
  data.table() %>% 
  separate(VS, c('Comparison','Reference'),' vs ') %>% 
  select(Reference, Comparison, logFC, Count) %>% 
  mutate(Count=as.numeric(Count)) %>% 
  arrange(Reference, Comparison)

all_counts

ggplot(all_counts, aes(x=paste(Comparison, Reference, sep = ' vs '), y=Count, fill=logFC)) + 
  geom_bar(stat='identity', position='dodge') + 
  theme_Publication() + 
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab('')
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

Set up gene lists for UpSet visualization
```{r, message=FALSE}
library(UpSetR)

fdr_005 <- list()
for (i in seq(1,length(comparisons_p))){
  genes <- topTable(efit, coef=i, adjust='fdr', number=30000, p.value=0.05) %>% rownames_to_column('Gene') %>% .[['Gene']] %>% list()
  fdr_005[comparisons_p[i]] <- genes
}

comparisons_p
```



UpSet of Retina <-> RPE | Cell Line <-> Adult
```{r, message = FALSE}
upset(fromList(fdr_005[c(1,6,7,9)]))
```

Splitting by up / down regulated
```{r, message=FALSE}
fdr_005_up <- list()
for (i in seq(1,length(comparisons_p))){
  genes <- topTable(efit, coef=i, adjust='fdr', number=30000, p.value=0.05) %>% filter(logFC > 1) %>% rownames_to_column('Gene') %>% .[['Gene']] %>% list()
  fdr_005_up[comparisons_p[i]] <- genes
}
fdr_005_down <- list()
for (i in seq(1,length(comparisons_p))){
  genes <- topTable(efit, coef=i, adjust='fdr', number=30000, p.value=0.05) %>% filter(logFC < -1) %>% rownames_to_column('Gene') %>% .[['Gene']] %>% list()
  fdr_005_down[comparisons_p[i]] <- genes
}

```


UpSet of Retina <-> RPE | Cell Line <-> Adult
Up and Down regualted in separate plots
```{r, message = FALSE}
upset(fromList(fdr_005_up[c(1,6,7,9)]),nintersects = NA, matrix.color = '#7fc97f', mainbar.y.max=1500)
upset(fromList(fdr_005_down[c(1,6,7,9)]), nintersects = NA, matrix.color = '#ef3b2c',mainbar.y.max=1500)
```



















UpSet of RPE (cell-line (ESC), cell-line(hTERT), fetal and adult tissue)
```{r, message = FALSE}
upset(fromList(fdr_005[1:5]),order.by = 'freq')
```

UpSet of RPE <-> Cornea <-> Retina
```{r, message = FALSE}
upset(fromList(fdr_005[c(6,11,12)]))
```

UpSet of ESC <-> RPE <-> Retina Cell-Lines
```{r, message = FALSE}
upset(fromList(fdr_005[c(7,8,10)]))
```

