---
title: "Figure 3"
output:
  html_notebook: default
  word_document: default
---
```{r,message = FALSE}
library(broom)
library(UpSetR)
library(tidyverse)
library(limma)
library(DT)
library(ReporteRs)
load('~/git/unified_gene_expression/data/big_six_DE_correct.Rdata')
load('~/git/unified_gene_expression/data/limma_voom_DE_all_by_all.Rdata')
source('~/git/unified_gene_expression/scripts/GO_enrichment.R')
source('~/git/unified_gene_expression/scripts/GO_term_finder.R')
# for GO enrichement, big six
background_genes <- topTable(big_six[[1]], number=30000) %>% rownames_to_column('Gene') %>% dplyr::select(Gene)
# for GO enrichment, all by all
background_genes_all_by_all <- topTable(efit_all, number=30000)%>% rownames_to_column('Gene') %>% dplyr::select(Gene)
options("ReporteRs-fontsize"=8, "ReporteRs-default-font"="Monaco")
```


Function to pull different sets out
```{r, message = FALSE}
set_maker2 <- function(list.of.all){
  out = list()
  # returns setdiff of list(s) against all lists
  # example: length(setdiff(intersect(test1_comps[[1]],test1_comps[[6]]),unlist(test1_comps[c(2,3,4,5)])))
  combinations <- combn(c(1,2,3,4,5,6),2)
  for (i in 1:6){
    print(names(list.of.all)[i])
    set2 <- setdiff(1:6, i)
    out[[names(list.of.all)[i]]] <- setdiff(list.of.all[[i]], unlist(list.of.all[set2])) 
  }
  for (i in 1:ncol(combinations)){
    set1 <- combinations[, i]
    print(paste(names(list.of.all)[set1], collapse = ', '))
    set2 <- setdiff(1:6, set1)
    out[[paste(names(list.of.all)[set1], collapse = ', ')]] <- setdiff(intersect(list.of.all[[set1[1]]], list.of.all[[set1[2]]]), unlist(list.of.all[set2]))
  }  
  out
}
```

Function to return gene list that are abs(logFC) > 1 and adj.P.Val < 0.05 for the different contrasts
```{r, messag= FALSE}
# just up vs 'control' (right side of contrast)
contrast_UPgene_lists <- function(limma_data, logFC){
  vector_of_comparisons <- colnames(limma_data)
  out_list <- list()
  for (i in 1:length(vector_of_comparisons)){
    stats <-  topTable(limma_data, coef=i, number = 30000, adjust.method = 'fdr', p.value = 0.05)
    if(nrow(stats)==0){
      out_list[vector_of_comparisons[i]] <- list('')
      next}
    stats_cut_down <- stats[stats[,'logFC'] > logFC,]
    out_list[vector_of_comparisons[i]] <- list(row.names(stats_cut_down))
    }
  out_list
}
# just down vs 'control' 
contrast_DOWNgene_lists <- function(limma_data, logFC){
  vector_of_comparisons <- colnames(limma_data)
  out_list <- list()
  for (i in 1:length(vector_of_comparisons)){
    stats <-  topTable(limma_data, coef=i, number = 30000, adjust.method = 'fdr', p.value = 0.05)
    if(nrow(stats)==0){
      out_list[vector_of_comparisons[i]] <- list('')
      next}
    stats_cut_down <- stats[stats[,'logFC'] < -logFC,]
    out_list[vector_of_comparisons[i]] <- list(row.names(stats_cut_down))
    }
  out_list
}
```


GO printer
```{r}
#GO printer
go_printer <- function(go_frame, num_to_print, title){
  out <- go_frame %>% arrange(as.numeric(`P value`)) %>% 
    head(n=num_to_print)  %>%  FlexTable(body.cell.props = cellProperties( padding = 2 ), 
                        header.par.props = parProperties(text.align = "left" ), 
                        body.par.props = parProperties(text.align = 'left'), header.columns = FALSE) %>%
    addHeaderRow(value = c(title), colspan = c(8), par.properties = parProperties(text.align='left')) %>% 
    addHeaderRow(value = colnames(go_frame)) %>% 
    setFlexTableWidths(widths = c(1, 1.4, 1.2, 1, 0.5, 0.7, 0.7, 6) )  %>% 
    setFlexTableBorders(inner.vertical = no_border, inner.horizontal = no_border,
                      outer.vertical = no_border, outer.horizontal = big_border )
  print(out)
}

# alt go printer to try and make a bit more sense out of the 80 bajillion results we get for retina vs rpe
go_printer2 <- function(go_frame, num_to_print, title){
  out <- go_frame %>% filter(as.numeric(`P value (FDR)`) < 0.05, as.numeric(Count) > 9, `Odds Ratio` > 4) %>% 
    head(n=num_to_print) %>% 
    FlexTable(body.cell.props = cellProperties( padding = 2 ), 
                        header.par.props = parProperties(text.align = "left" ), 
                        body.par.props = parProperties(text.align = 'left'), header.columns = FALSE) %>%
    addHeaderRow(value = c(title), colspan = c(8), par.properties = parProperties(text.align='left')) %>% 
    addHeaderRow(value = colnames(go_frame)) %>% 
    setFlexTableWidths(widths = c(1, 1.4, 1.2, 1, 0.5, 0.7, 0.7, 6) )  %>% 
    setFlexTableBorders(inner.vertical = no_border, inner.horizontal = no_border,
                      outer.vertical = no_border, outer.horizontal = big_border )
  print(out)
}
```

All GO comparison runner. Takes in the output from the set generator and makes does GO enrichment for all 6 unique categories (a, b, c, ab, bc, abc)
```{r, message = FALSE}
all_go_tester <- function(test_set, test_name, background_genes){
  column_classes <- rep(x='character',11)
  column_names <- c("GO ID","P value","P value (FDR)","Odds Ratio","Expected Count","Count","Size","Term","Test", "Set", "Ontology")
  empty_frame <- read.table(text='', colClasses = column_classes, col.names = column_names)
  all_go_terms = empty_frame
 
  for (i in names(test_set)){
    if (length(test_set[[i]])<6){
      print(paste(i, 'has less than 6 genes in it, skipping GO enrichment'))
    } 
    else {
      print(i)
      for (ontology in c('BP','MF')){
        go_run <- GO_enrichment(test_set[[i]], background_genes, ontology)
        go_run <- go_run %>% data.frame() %>% mutate(Test=test_name, Set=i, Ontology=ontology)
        colnames(go_run)[1] <- 'GO.ID'
        print(colnames(go_run)==colnames(all_go_terms))
        all_go_terms <- rbind(all_go_terms, go_run)
      }
    }
  }
  
colnames(all_go_terms) <- c("GO ID","P value","P value (FDR)","Odds Ratio","Expected Count","Count","Size","Term","Test", "Set", "Ontology")
all_go_terms
}
```

Six DE tests:

1. Cornea vs Retina vs RPE (all adult)
2. ESC vs RPE vs Retina (all Stem cell lines, non immortalized)
3. Adult RPE vs Fetal RPE vs Cell Line RPE
4. immortalized RPE vs Stem Cell Line RPE
5. Adult Retina vs Stem Cell Line Retina
6. Adult Cornea vs Fetal Cornea vs Cell Line Cornea


# 1. Cornea vs Retina vs RPE (all adult)
```{r, message = FALSE, fig.width=3, fig.height=3}
testA_go_all_sets <- all_go_tester(set_maker2(test1_comps),'Test A', background_genes)

#go_printer(testA_go_all_sets %>% filter(Test=='Test A',Set=='b',Direction=='Down') %>%  dplyr::select(`GO BP ID`:Term),15, 'Test A - Set B Down')
```


# 2. ESC vs RPE vs Retina (all Stem cell lines, non immortalized)
```{r, message = FALSE}
testB_go_all_sets <- all_go_tester(set_maker2(test2_comps),'Test B', background_genes)
```

# 3. Adult RPE vs Fetal RPE vs Cell Line RPE
```{r}
testC_go_all_sets <- all_go_tester(set_maker2(test3_comps),'Test C', background_genes)

#go_printer(testC_go_all_sets %>% filter(Test=='Test C',Set=='b',Direction=='Up',as.numeric(`P value (FDR)`)<0.05) %>%  dplyr::select(`GO BP ID`:Term),15, 'Test C - Set B Down')
```

# 4. immortalized RPE vs Stem Cell Line RPE
```{r, messager = FALSE}
colnames(big_six[[4]])
test4_contrast_genes_up <-contrast_UPgene_lists(big_six[[4]], 1)
test4_contrast_genes_down <-contrast_DOWNgene_lists(big_six[[4]], 1)
testD_go_all_sets <- rbind(
                            GO_enrichment(test4_contrast_genes_up, background_genes, 'BP') %>% 
                              mutate(Test='Test D', Set='RPE (cell) > RPE (stem cell)', Ontology='BP') %>% 
                              rename(`GO BP ID`='GO ID'),
                           GO_enrichment(test4_contrast_genes_up, background_genes, 'MF') %>% 
                              mutate(Test='Test D', Set='RPE (cell) > RPE (stem cell)', Ontology='MF') %>% 
                              rename(`GO MF ID`='GO ID'),
                           GO_enrichment(test4_contrast_genes_down, background_genes, 'BP') %>% 
                              mutate(Test='Test D', Set='RPE (cell) < RPE (stem cell)', Ontology='BP') %>% 
                              rename(`GO BP ID`='GO ID'),
                            GO_enrichment(test4_contrast_genes_down, background_genes, 'MF') %>% 
                              mutate(Test='Test D', Set='RPE (cell) < RPE (stem cell)', Ontology='MF') %>% 
                              rename(`GO MF ID`='GO ID')
                           )
```
# 5. Adult Retina vs Stem Cell Line Retina
```{r, message =  FALSE}
colnames(big_six[[5]])
test5_contrast_genes_up <-contrast_UPgene_lists(big_six[[5]], 1)
test5_contrast_genes_down <-contrast_DOWNgene_lists(big_six[[5]], 1)
testE_go_all_sets <- rbind(
                            GO_enrichment(test5_contrast_genes_up, background_genes, 'BP') %>% 
                              mutate(Test='Test E', Set='Retina > Retina (stem cell)', Ontology='BP') %>% 
                              rename(`GO BP ID`='GO ID'),
                           GO_enrichment(test5_contrast_genes_up, background_genes, 'MF') %>% 
                              mutate(Test='Test E', Set='Retina > Retina (stem cell)', Ontology='MF') %>% 
                              rename(`GO MF ID`='GO ID'),
                           GO_enrichment(test5_contrast_genes_down, background_genes, 'BP') %>% 
                              mutate(Test='Test E', Set='Retina < Retina (stem cell)', Ontology='BP') %>% 
                              rename(`GO BP ID`='GO ID'),
                            GO_enrichment(test5_contrast_genes_down, background_genes, 'MF') %>% 
                              mutate(Test='Test E', Set='Retina < Retina (stem cell)', Ontology='MF') %>% 
                              rename(`GO MF ID`='GO ID')
                           )
```


# 6. Adult Cornea vs Fetal Cornea vs Cell Line Cornea
```{r, message = FALSE}
testF_go_all_sets <- all_go_tester(set_maker2(test6_comps),'Test F', background_genes)
```

# How to find what genes are the GO:ID
```{r, message = FALSE}
#go_associated_genes <- GO_term_finder(retina_rpe_GO_genes_up,'GO:0050953')
#topTable(big_six[[1]],number=30000) %>% rownames_to_column('Gene') %>% filter(Gene %in% go_associated_genes)
```

# Save GO term enrichment searches
```{r, message = FALSE}
go_enrichment <- bind_rows(testA_go_all_sets,
                       testB_go_all_sets,
                       testC_go_all_sets,
                       testD_go_all_sets,
                       testE_go_all_sets,
                       testF_go_all_sets)
save(go_enrichment, file='~/git/unified_gene_expression/data/go_enrichment.Rdata')
```

# Summarize num of significant results (GO)
```{r, message = FALSE}
go_enrichment %>% group_by(Test, Set, Ontology)  %>% summarise(Count=sum(as.numeric(`P value (FDR)`) < 0.05))
```

# Print out GO results as csv
```{r, message = FALSE}
out_csv = data.frame()
for (i in names(go_enrichment)) {
  out <- go_enrichment[[i]] %>% mutate(Test = gsub('\\.|_',' ', i)) %>% head(n=15)
  out_csv <- rbind(out_csv, out)
}
#write.csv(out_csv, file='~/Dropbox/Work/NEI/PROJECTS/mcgaughey/eyeIntegration/supplemental_table_s3.csv')

out_csv = data.frame()
for (i in names(go_enrichment)) {
  sig_results = go_enrichment[[i]] %>% filter(as.numeric(`P value`)<0.01) %>% nrow()
  results_to_keep = max(sig_results, 10)
  out <- go_enrichment[[i]] %>% mutate(Test = gsub('\\.|_',' ', i)) %>% head(n=results_to_keep)
  out_csv <- rbind(out_csv, out)
}

#write.csv(out_csv, file='~/Dropbox/Work/NEI/PROJECTS/mcgaughey/eyeIntegration/supplemental_table_s4.csv')
```


#################
# all by all GO
#################

Generate GO enrichment for eye vs body
```{r, message = FALSE}
up_gene_lists <- contrast_UPgene_lists(efit_all, 2)
# only keep body comparisons
# up_gene_lists <- up_gene_lists[grep('Body',names(contrast_UPgene_lists(efit_all, 2)))]

down_gene_lists <- contrast_DOWNgene_lists(efit_all, 2)
# only keep body comparisons
# down_gene_lists <- down_gene_lists[grep('Body',names(contrast_DOWNgene_lists(efit_all, 2)))]

go_up <- all_go_tester(up_gene_lists,'Up', background_genes_all_by_all)
go_down <- all_go_tester(down_gene_lists,'Down', background_genes_all_by_all)

all_vs_all_go <- rbind(go_up, go_down)

all_vs_all_go %>% mutate(Set = ifelse(Test=='Down', gsub('_vs_',' < ', Set),gsub('_vs_',' > ', Set) )) %>% group_by(Set, Test, Ontology) %>% filter(as.numeric(`P value (FDR)`)<0.01) %>%  summarise(Count=n())
```

RPE vs Retina adult __ VS __ RPE vs Retina stem
```{r}

# ## adult
# test1_contrast_genes <- contrast_UPgene_lists(big_six[[1]])
# sets <- set_maker(test1_contrast_genes$Cornea.adult_vs_Retina.adult,
#                   test1_contrast_genes$Cornea.adult_vs_RPE.adult,
#                   test1_contrast_genes$Retina.adult_vs_RPE.adult)
# retina.adult_vs_rpe.adult_up_genes <- sets$c
# 
# test1_contrast_genes <- contrast_DOWNgene_lists(big_six[[1]])
# sets <- set_maker(test1_contrast_genes$Cornea.adult_vs_Retina.adult,
#                   test1_contrast_genes$Cornea.adult_vs_RPE.adult,
#                   test1_contrast_genes$Retina.adult_vs_RPE.adult)
# retina.adult_vs_rpe.adult_down_genes <- sets$c
# 
# ## stem cells
# test2_contrast_genes <- contrast_UPgene_lists(big_six[[2]])
# sets <- set_maker(test2_contrast_genes$ESC.stemcell_vs_Retina.stemcell,
#                   test2_contrast_genes$ESC.stemcell_vs_RPE.stemcell,
#                   test2_contrast_genes$Retina.stemcell_vs_RPE.stemcell)
# retina.stem_vs_rpe.stem_up_genes <- sets$c
# 
# test2_contrast_genes <- contrast_DOWNgene_lists(big_six[[2]])
# sets <- set_maker(test2_contrast_genes$ESC.stemcell_vs_Retina.stemcell,
#                   test2_contrast_genes$ESC.stemcell_vs_RPE.stemcell,
#                   test2_contrast_genes$Retina.stemcell_vs_RPE.stemcell)
# retina.stem_vs_rpe.stem_down_genes <- sets$c
# 
# 
# 
# ## set operations
# length(retina.adult_vs_rpe.adult_up_genes)
# length(retina.adult_vs_rpe.adult_down_genes)
# length(retina.stem_vs_rpe.stem_up_genes)
# length(retina.stem_vs_rpe.stem_down_genes)
# length(intersect(retina.adult_vs_rpe.adult_up_genes, retina.stem_vs_rpe.stem_up_genes))
# length(intersect(retina.adult_vs_rpe.adult_down_genes, retina.stem_vs_rpe.stem_down_genes))

# let's find genes that are up in retina stem cells (relative to rpe stem) but NOT up
```